Using cosmic variance to constrain the dark matter halo mass of Lyman-alpha emitting galaxies at $z$=3.1

### 2013-03-20 paper/ms.tex
new figure, updated text on obs contraints for the occupation fraction

\newcommand{\apj}{ApJ}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}

@@

\title[Dark Matter Halo Mass for LAEs  at $z=3.1$]{Constraints on the
  dark matter halo mass of Lyman-alpha emitting galaxies at $z$=3.1
  from cosmic variance} 
\author[J.E. Forero-Romero and J. Mejia]{
\parbox[t]{\textwidth}{\raggedright 
  Julian Mejia$^{2}$ and
Jaime E. Forero-Romero$^{1}$}
\vspace*{6pt}\\
$^{1}$\\
$^{2}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1 No. 18A-10, Edificio Ip, Bogot\'a, Colombia
}

@@

\begin{abstract}
In this Letter we implement a new method to find the characteristic
mass of dark matter halos hosting Lyman-Alpha Emitting (LAE)
galaxies. The method is based on matching the statistics for the
cosmic variance with the observations recently obtained by
\cite{Yamada2012}. We build a model for LAEs on top of a large N-body
cosmological simulation (Bolshoi). The basic principle is that a dark
matter halo can host one LAE with a probability $f_{\rm occ}$ if its
mass is found withing a certain range mass range delimited by two
threshold values, $M_{\rm min}$ and $M_{\rm max}$. We generate
hundreds of mock catalogs in a thourough exploration of the parameter
space of this model. These catalogs are constructed in such a way as
to reproduce the spatial correlation between the different observed
fields. We find that the models that best match the observed cosmic
variance statistics are those with halo masses in the range $M_{\rm
  min}=$ and $M_{\rm max}=$ with and occupation fraction that scales
as $f_{\rm occ}=$ with the minimum mass. We find that the angular
correlation function is in agreement with the observational
constraints mostly due to the cosmic variance in the small angular
fields where this statistics has been computed so far. We also present
predictions for future measurements of the correlation function on
larger continous fields. We make available the mock data for the best
models in a public repository.  
\end{abstract}

@@

Throughout this Letter we assume a $\Lambda$CDM cosmology with the
following values for the cosmological parameters, $\Omega_{m}=0.27$,
$\Omega_{\Lambda}=0.73$ and $h=0.70$, corresponding to the matter
density, vacuum density and the Hubble constant in units of 100 km
s$^{-1}$ Mpc$^{-1}$.

\section{Methodology}
In this Letter we constrain the typical mass of dark matter halos
hosting LAES at $z=3.1$. Our model is based on the number density
information obtained in the recent large scale survey presented by XXX
where XXX LAEs are detected over 7 fields of $\sim 46 \times
35$Mpc$^{2}h^{-2}$ in area comoving in area corresponding to observed
fields of $XXX$ deg$^{2}$.

@@

In this Letter we want to step back and reduce the complexity of our
model, with the sole objective of reproducing the cosmic variance in
the number density of LAEs. Afterwards we will interpret the
implications of this result for physical models for Lyman-alpha
emitting galaxies.

@@

Our observational reference are the recently published results of a
panoramic survey of LAES at z=3.1 by \cite{Yamada2012}. This survey
was conducted with the Subaru 8.2m telescope and the Subaru Prime
Focus Camera, which has a field of view covering $34\times 27$ arcmin,
corresponding to a comoving scale of $46\times35$ Mpc $h^{-1}$ at
$z=3.09$. The narrow band filter is centered at $4977$ \AA with a
$77$\AA width, corresponding to the redshift range $z=3.062-3.125$ and
$41$ Mpc $h^{-1}$ comoving scale for the detection of the
Lyman-$\alpha$ line centered at $z=3.09$.

@@

We use halo catalogs constructed with a Friend-of-Friends (FOF)
algorithm with a linking lenght of 0.17 times the interparticle
distance. We have veryfied that the main results we present in this
paper also hold if instead we use halo catalogs constructed from a the
Bound Density Maxima (BDM) algorithm \citep{KlypinBDM} that are
defined to have an density of 200 times the critical density. The
minimum halo mass in the models we construct in this Letter correspond
to groups of $\sim 75$ particles. The catalogs were obtained from the
publicly available Multidark database \footnote{{\tt
    http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}.

@@

\subsection{Dark Matter Halo Number Density}
In Figure \ref{fig:laes_dist} we present the results for  the intergrated dark matter halo surface
density as a function of halos mass for each one of the 210 sub-volumes in
the Bolshoi simulation. The continous lines correspond to the
subvolumes. The shadowed area indicates the surface density values for
LAEs allowed by the observations.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig2_match_P5.pdf}
\hspace{5mm}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig2_match_P99.pdf}\\
\vspace{5mm}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig2_random_P5.pdf}
\hspace{5mm}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig2_full_P5.pdf}
\end{center} 
\caption{$M_{\rm min}$-$M_{\rm max}$ plane for all models with
  $P>0.05$\label{figure:landscape}.} 
\end{figure*}

@@

Figure \label{figure:landscape} presents the regions in the parameter
space $M_{\rm min}-M_{\rm max}$ where the KS test yields values of
$P>0.1$. Each panel corresponds to the three different ways of
grouping the mock fields. In the case of the methods {\tt{Match}}
and {\tt{Random}} the color code indicates the fraction of these 15
mock surveys with with $P>0.1$. The third panel shows the result for
the method {\tt{Full}}, in this case the color code correspond
to the value of $P$ for a given model.

These results distinguish three mass regimes. In the first regime, at
high mass values, we find that LAE models with minimum mass of $M_{\rm
  min}>3\times 10^{11}$\hMsun are not compatible with observations as
expected from the analysis in the previous subsection. For masses below $M_{\rm
  min}=3\times 10^{10}$ any values for $M_{\rm min}$ and $M_{\rm max}$
can be made compatible with observations, provided that
$f_{\rm occ}$ is fine tuned to do it. In an intermediate mass regime,
for minimum mass values $3\times 10^{10}\hMsun < M_{\rm min}< 3\times 10^{11}\hMsun$ only a
limited range of models with $M_{\rm max}$ with occupation fraction
$f_{\rm occ}\sim 1$ is able to reproduce observations.

@@

@@

\section{Conclusions}
In this Letter we have constrained the preferred mass for dark matter
halos hosting Lyman Alpha Emitters at a redshift $z=3.1$. We use a
method that matches the cosmic variance in the surface
density number of LAEs from differente observed fields. From this we
are able to constrain the mass range of dark matter halos hosting LAEs
to be in the range $<M_{\rm h}<$ and a corresponding occupation fraction
that escales as $f_{\rm occ} = M_{\rm min}$ with the minium halo mass
defining the halo mass range.

@@

>>>>

\newcommand{\documentname}{paper~}
\newcommand{\apj}{ApJ}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}

@@

\title[Dark Matter Halo Mass for LAEs  at $z=3.1$]{Using cosmic
  variance to constrain the
  dark matter halo mass of Lyman-alpha emitting galaxies at $z$=3.1}
\author[~J.~E. Forero-Romero and J. Mejia]{
\parbox[t]{\textwidth}{\raggedright 
  Julian Mej\'ia$^{2}$ and
Jaime E. Forero-Romero$^{1}$}
\vspace*{6pt}\\
$^{1}$\\
$^{2}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia 
}

@@

\begin{abstract}
We use simulations and recent observations to find the characteristic
mass of dark matter halos hosting Lyman-Alpha Emitting (LAE) galaxies
at a redshift of $z=3.1$. The method is based on matching the
statistics for the number density cosmic variance between different
observed and mock fields. We build a model for LAEs on top of a large N-body
cosmological simulation (Bolshoi) where the basic principle is that a dark
matter halo can only host one LAE with a probability $f_{\rm occ}$ if its
mass is found withing a certain range mass range delimited by two
threshold values, $M_{\rm min}$ and $M_{\rm max}$. We generate
hundreds of mock catalogs in a thourough exploration of the parameter
space of this model. These catalogs are constructed in such a way as
to reproduce the spatial correlation between the different observed
fields. We find that the models that best match the observed cosmic
variance statistics are those with halo masses in the range $M_{\rm
  min}=$ and $M_{\rm max}=$ with and occupation fraction that scales
as $f_{\rm occ}=$ with the minimum mass. We find that the angular
correlation function is in agreement with the observational
constraints mostly due to the cosmic variance in the small angular
fields where this statistics has been computed so far. We also present
predictions for future measurements of the correlation function on
larger continous fields. We make available the mock data for the best
models in a public repository.  
\end{abstract}

@@

Throughout this \documentname we assume a $\Lambda$CDM cosmology with the
following values for the cosmological parameters, $\Omega_{m}=0.27$,
$\Omega_{\Lambda}=0.73$ and $h=0.70$, corresponding to the matter
density, vacuum density and the Hubble constant in units of 100 km
s$^{-1}$ Mpc$^{-1}$.

\section{Methodology}
In this \documentname we constrain the typical mass of dark matter halos
hosting LAES at $z=3.1$. Our model is based on the number density
information obtained in the recent large scale survey presented by XXX
where XXX LAEs are detected over 7 fields of $\sim 46 \times
35$Mpc$^{2}h^{-2}$ in area comoving in area corresponding to observed
fields of $XXX$ deg$^{2}$.

@@

In this \documentname we want to step back and reduce the complexity of our
model, with the sole objective of reproducing the cosmic variance in
the number density of LAEs. Afterwards we will interpret the
implications of this result for physical models for Lyman-alpha
emitting galaxies.

@@

Our observational reference are the recently published results of a
panoramic survey of LAES at $z=3.1$ by \cite{Yamada2012}. This survey
was conducted with the Subaru 8.2m telescope and the Subaru Prime
Focus Camera, which has a field of view covering $34\times 27$ arcmin,
corresponding to a comoving scale of $46\times35$ Mpc $h^{-1}$ at
$z=3.09$. The narrow band filter is centered at $4977$ \AA with a
$77$\AA width, corresponding to the redshift range $z=3.062-3.125$ and
$41$ Mpc $h^{-1}$ comoving scale for the detection of the
Lyman-$\alpha$ line centered at $z=3.09$.

@@

We use halo catalogs constructed with a Friend-of-Friends (FOF)
algorithm with a linking lenght of 0.17 times the interparticle
distance. We have veryfied that the main results we present in this
paper also hold if instead we use halo catalogs constructed from a the
Bound Density Maxima (BDM) algorithm \citep{KlypinBDM} that are
defined to have an density of 200 times the critical density. The
minimum halo mass in the models we construct in this \documentname correspond
to groups of $\sim 75$ particles. The catalogs were obtained from the
publicly available Multidark database \footnote{{\tt
    http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}.

@@

\subsection{Dark Matter Halo Number Density}
In Figure \ref{fig:laes_dist} we present the results for  the
integrated dark matter halo surface density as a function of halos
mass. Each line correspond to one of the 210 sub-volumes in the
Bolshoi simulation. The shadowed area indicates the surface density
values for LAEs allowed by the observations.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.95\linewidth,angle=0]{./plots/Fig2_match_P5.pdf}\\
\vspace{5mm}
\includegraphics[width=0.95\linewidth,angle=0]{./plots/Fig2_random_P5.pdf}\\
\vspace{5mm}
\includegraphics[width=0.95\linewidth,angle=0]{./plots/Fig2_full_P5.pdf}
\end{center} 
\caption{$M_{\rm min}$-$M_{\rm max}$ plane for all models with
  $P>0.05$ in the three different of grouping the mock fields. In the
  case of the {\tt match} and {\tt random} methods the  color code
  corresponds to the number of mock surveys that are found to
  be compatible with observations. For the {\tt full} method the
  colour code corresponds to the results of $100$ times the maximum $P$ value
  resulting from the KS test. \label{figure:landscape}.}  
\end{figure}

@@

Figure \label{figure:landscape} presents the regions in the parameter
space $M_{\rm min}$-$M_{\rm max}$ where the KS test yields values of
$P>0.05$. We consider that for those models it is not possible to rule out the null
hypothesis, namely that the number density in simulated data and the observations come
from the same parent distribution. Each panel corresponds to the three different ways of
grouping the mock fields. In the case of the methods {\tt{Match}}
and {\tt{Random}} the color code indicates the fraction of these 15
mock surveys with with $P>0.1$. The third panel shows the result for
the method {\tt{Full}}, in this case the color code correspond
to the value maximum value of $100\times P$ for a model with those
mass ranges.

These results clearly distinguish three mass regimes. In the first regime, at
high mass values, we find that LAE models with minimum mass of $M_{\rm
  min}>10^{11.5}$\hMsun are not compatible with observations. There is
a second regime for masses below $M_{\rm
  min}=3\times 10^{10}$ any values for $M_{\rm min}$ and $M_{\rm max}$
can be made compatible with observations, provided that
$f_{\rm occ}$ is fine tuned to do it. In an intermediate mass regime,
for minimum mass values $3\times 10^{10}\hMsun < M_{\rm min}< 3\times 10^{11}\hMsun$ only a
limited range of models with $M_{\rm max}$ with occupation fraction
$f_{\rm occ}\sim 1$ is able to reproduce observations.

@@

\subsection{Additional observational constraints on the occupation fraction}

\cite{Hayes2010} constrained the value of $f_{\rm occ}$ at $z=2.2$ to
be $f_{\rm occ}=0.10$. This estimation was based on blind surveys of
the H$\alpha$ and Lyman alpha line with the European Southern
Observatorio (ESO) Very Large Telescope (VLT). Using corrections by
extinction to obtain an estimate for the intrinsic H$\alpha$
luminosity, and using values for the theoretical expectation of the
ratio Lyman$\alpha$/H$\alpha$ they derive an bulk escape fraction for
the Lyman$\alpha$ radiation of $f_{\rm esc}=(5.3\pm 3.8)\%$ or $f_{\rm
esc}=(10.7\pm 2.8)$ if a different dust correction is used. From this
results they derive that almost $90\%$ of the star forming galaxies
emit insufficient Lyman-alpha to be detected, effectively setting the
occupation fraction to be $f_{\rm occ}=0.10$. Obervations are
consistent with these numbers are
consistent with being constant for every galaxy regardless of its
luminosity.

For the cosmological parameters used in this \documentname the age of
the universe between $z=3.1$ and $z=2.2$ has changed by $\sim 1$ Gyr,
it is reasonable to assume that the physical conditions that determine
the escape fraction $f_{\rm esc}$ and the occupation fraction $f_{\rm
  occ}$ remain constant. This assumption allows us to further pick
models that have an occupation fraction of $f_{\rm occ}=0.10$. Under
this selection only $18$ models can be selected. Considering an
occupation fraction $f_{\rm occ}=0.20$ another $39$ models can be
considered. The minimum and maximum mass are found in the appendix in
Table XX.

@@

\section{Conclusions}
In this \documentname we have constrained the preferred mass for dark matter
halos hosting Lyman Alpha Emitters at a redshift $z=3.1$. We use a
method that matches the cosmic variance in the surface
density number of LAEs from differente observed fields. From this we
are able to constrain the mass range of dark matter halos hosting LAEs
to be in the range $<M_{\rm h}<$ and a corresponding occupation fraction
that escales as $f_{\rm occ} = M_{\rm min}$ with the minium halo mass
defining the halo mass range.

@@

\begin{table}
\begin{center}
\begin{tabular}{ccc}\hline\hline
10.1 & 10.2 & 0.1\\
10.3 & 10.5 & 0.1\\
10.4 & 10.7 & 0.1\\
10.5 & 10.9 & 0.1\\
10.6 & 11.2 & 0.1\\
10.6 & 11.3 & 0.1\\
10.6 & 11.4 & 0.1\\
10.1 & 10.2 & 0.1\\
10.3 & 10.5 & 0.1\\
10.4 & 10.7 & 0.1\\
10.5 & 10.9 & 0.1\\
10.5 & 11.0 & 0.1\\
10.6 & 11.2 & 0.1\\
10.6 & 11.3 & 0.1\\
10.6 & 11.4 & 0.1\\
10.6 & 11.5 & 0.1\\
10.6 & 11.6 & 0.1\\
10.6 & 11.7 & 0.1\\\hline\hline

10.4 & 10.5 & 0.2\\
10.6 & 10.8 & 0.2\\
10.7 & 11.0 & 0.2\\
10.8 & 11.2 & 0.2\\
10.8 & 11.3 & 0.2\\
10.9 & 11.6 & 0.2\\
10.9 & 11.7 & 0.2\\
10.9 & 11.8 & 0.2\\
10.9 & 11.9 & 0.2\\
10.9 & 12.0 & 0.2\\
10.9 & 12.1 & 0.2\\
10.9 & 12.2 & 0.2\\
10.9 & 12.3 & 0.2\\
10.9 & 12.4 & 0.2\\
10.9 & 12.5 & 0.2\\
10.9 & 12.6 & 0.2\\
10.9 & 12.7 & 0.2\\
10.9 & 12.8 & 0.2\\
10.9 & 12.9 & 0.2\\
10.9 & 13.0 & 0.2\\
10.6 & 10.8 & 0.2\\
10.7 & 11.0 & 0.2\\
10.8 & 11.2 & 0.2\\
10.8 & 11.3 & 0.2\\
10.8 & 11.4 & 0.2\\
10.9 & 11.7 & 0.2\\
10.9 & 11.8 & 0.2\\
10.9 & 11.9 & 0.2\\
10.9 & 12.0 & 0.2\\
10.9 & 12.1 & 0.2\\
10.9 & 12.2 & 0.2\\
10.9 & 12.3 & 0.2\\
10.9 & 12.4 & 0.2\\
10.9 & 12.5 & 0.2\\
10.9 & 12.6 & 0.2\\
10.9 & 12.7 & 0.2\\
10.9 & 12.8 & 0.2\\
10.9 & 12.9 & 0.2\\
10.9 & 13.0 & 0.2\\\hline\hline
\end{tabular}
\end{center}
\end{table}
